Short-term hourly forecasting for care services

### 2020-06-02 paper/paper.Rmd
add introduction and new graphs to data subsection

---
title: Short-term hourly forecasting in urgent and emergency care
author:
  - name: Author1
    email: email1@example.com
    affiliation: University1
    footnote: 1
  - name: Author2
    email: email2@example.com
    affiliation: University2
    footnote: 2
  - name: Author3
    email: email3@example.com
    affiliation: University3
    footnote: 2
address:
  - code: University1
    address: Cardiff business school, 3 Colum Drive, CF10 3EU, Cardiff
  - code: University2
    address: adress2
  - code: University3
    address: adress3
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"

@@

journal: "Which journal? Journal of Service Research/Health Services Research/EJOR/IJF"
date: "`r Sys.Date()`"
geometry: "top=25mm, left=30mm, right=30mm, bottom=25mm,headsep=10mm, footskip=12mm"
#linenumbers: true
numbersections: true
#output: rticles::elsevier_article
header-includes:  
 \usepackage{adjustbox, float} #use the 'float' package
output:
  bookdown::pdf_book: 
    base_format: rticles::elsevier_article
    citation_package: natbib
bibliography: mybibfile.bib
biblio-style: authoryear
#csl: elsevier-without-titles.csl
#biblio-style: abbrevnat
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, fig.pos = 'H')
# Make sure you have the latest versions of rmarkdown and rticle
library(tidyverse)
library(lubridate)
library(bookdown)
h2_hourly <- readr::read_rds("../data/hw_hourly.rds")
```

@@

@@

1. We measure the characteristics of the time series at the original level and at various levels of aggregation. We reveal how TA influences the time series characteristics. 
2. We develope a model to discover the relationship between the original time series characteristics and the forecasting performance evaluation results to construct rules for method selection.
3. We determine for which type of time series characteristics, each TA forecasting strategy is preferable.

The rest of the paper is organised as follwoing: section \ref{lit} provides a brief overview of the use of temporal aggregation in time series forecasting. Section \ref{model} starts with ...

# Research background:  Hourly short-term forecasting in emergency care {#lit}

@@

\citet{hertzum2017forecasting} & Hourly ED patient arrivals and ED occupancy forecasting using calendar variables  1-3 & Regression, ARIMA, naive & MAPE & blabla\\ 
\citet{mccarthy2008challenge} & To develop methodology for predicting demand for ED services by characterizing ED arrivals  1-3 & Poissopn regression & RMSE & blabla\\ 
\citet{morzuch2006forecasting} & hourly ED arrivals for 24 hours & Holt-Winters & MAPE & blabla\\

@@

Linear regression, ARIMA, and naive models were used by @hertzum2017forecasting to investigate whether accurate hourly accident and emergency department patient arrivals and occupancy forecasts can be generated using calendar variables. Naive model was there for the purpose of comparison. @hertzum2017forecasting study shows that patient arrivals variation is larger across the hours of the day than across the days of the week and the months of the year. In term of hour of the day, patient arrivals peaked around noon. For days of the week, Monday is the busiest day while weekends are the quietest days. July-August are the month with the highest number of patient arrivals and January and February are the months with the lowest number of arrivals. The regression and ARIMA models perform similarly for all forecast interval in modeling patient arrivals. In modeling accident and emergency department occupancy, ARIMA outperform regression models. However, afterall, the models of occupancy were less accurate than those arrivals. @hertzum2017forecasting mentioned that ARIMA models are among the most accurate models for accident and emergency department visits forecasting. Another interesting point is that the accuracy of accident and emergency department forecasting models decrease with the increasing forecast interval. Lastly, the accuracy of the forecasting model may possibly be increased with additional information added to the model.

@@

```{r hourly-plot, fig.cap = 'Seasonal plot of ED attendance', fig.align='center',out.width="80%"}
h2_hourly <- h2_hourly %>% filter(arrival_1h >="2015-01-01 00:00:00"&arrival_1h <"2019-01-01 00:00:00")
  subseriesplot <- h2_hourly %>% as_tibble() %>% mutate(day=factor(yday(arrival_1h)),year=factor(year(arrival_1h)),hour=hour(arrival_1h)) %>% 
    select(-arrival_1h)
time <-  format( seq.POSIXt(as.POSIXct(Sys.Date()), as.POSIXct(Sys.Date()+1), by = "1 hour"),
                 "%H:%M", tz="GMT")
time <- time[-length(time)]
p1 <- ggplot(data=subseriesplot,mapping = aes(x=hour ,y=n_attendance, colour=day))+
    geom_line(aes(group=interaction(day,year)))+
    geom_point()+
    theme_bw()+
    theme(legend.position = "")+
    labs(y="Number of visits in ED", x="Hour of day")+
      scale_x_continuous(breaks=0:23, labels = time)+
  theme(axis.text.x = element_text(angle = 90))
  p1+theme(plot.title = element_text(size = 10))
```

@@

```{r 24hour, fig.cap= "Distrubution of ED attendance for each hour of day", fig.align='center',fig.width=6, fig.height=9}
h2_hourly %>% as_tibble() %>% mutate(Date=date(arrival_1h),Hour=hour(arrival_1h)) %>% 
  select(-arrival_1h) %>% ggplot(mapping=aes(x=n_attendance))+ 
geom_histogram(mapping=aes(y=stat(density)), binwidth = 1,fill="white",color = 'black')+
  geom_density(color="blue")+
scale_y_continuous(labels = scales::percent_format())+
facet_wrap(~Hour, scales = "free", ncol = 3)+
theme_bw()+
labs(x="Number of visits in ED", y="Percentage")+
theme(axis.text = element_text(size = 7))
```

>>>>

---
title: Short-term hourly forecasting for care services
author:
  - name: Author1
    email: email1@example.com
    affiliation: University1
    footnote: 1
  - name: Author2
    email: email2@example.com
    affiliation: University2
    footnote: 2
  - name: Author3
    email: email3@example.com
    affiliation: University3
    footnote: 2
address:
  - code: University1
    address: Cardiff business school, 3 Colum Drive, CF10 3EU, Cardiff
  - code: University2
    address: adress2
  - code: University3
    address: adress3
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"

@@

journal: "Which journal? Journal of Service Research/Health Services Research/EJOR/IJF"
date: "`r Sys.Date()`"
geometry: "top=25mm, left=30mm, right=30mm, bottom=25mm,headsep=10mm, footskip=12mm"
#linenumbers: true
numbersections: true
#output: rticles::elsevier_article
header-includes:  
 \usepackage{adjustbox, float,lscape} #use the 'float' package
output:
  bookdown::pdf_book: 
    base_format: rticles::elsevier_article
    citation_package: natbib
bibliography: mybibfile.bib
biblio-style: authoryear
#csl: elsevier-without-titles.csl
#biblio-style: abbrevnat
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, fig.pos = 'H')
# Make sure you have the latest versions of rmarkdown and rticle
library(tidyverse)
library(lubridate)
library(bookdown)
library(patchwork)
library(tsibble)
h2_hourly <- readr::read_rds("../data/hw_hourly.rds")
```

@@

why forecasting for urgent and emergency care is important?

Urgent and emergency healthcare is increasingly regarding patients as consumers of its service and aims to provide 'consumer satisfaction'; particularly in the Accident and Emergency (A&E) department and Ambulance services which have been viewed as the 'shop window' of the hospital service. An accurate forecasting of the demand is essential in A&E and Ambulance service trust to depict various courses of action that can result in massive savings in terms of patient lives.

<!-- Demand forecasting is a critical step to inform bed capacity planning and staffing to meet the needs of patients on each hospital service and inpatient unit. One of the best ways to optimize staffing utilization and related costs is to forecast demand far enough in advance in order to sufficiently match staff to meet anticipated patient demand without incurring last-minute expenses, such as overtime or supplemental staffing. -->
<!-- Accurate forecasting of patient demand by season, month, day of the week, and time of day enables leaders to making strategic decisions regarding the resources needed to provide services to the number and type of patients requiring hospital services.  -->
<!-- Equipped with accurate demand forecasts, hospital leaders can reconfigure units and redeploy staff during predicted low-census periods, and units with seasonal occupancy can be closed and repurposed. -->

Accurate forecasts contribute to better allocation of resources. Patient overcrowding in urgent and emergency care is a serious problem that causes challenging situations on patient flow . Also, it is related with increasing length of stay (LOS), low patient satisfaction about treatments on patients, number of patients left the ED without seen by medical staff , unexpected return visits to EDs , increasing health care costs and some other miscellaneous problems such as inaccuracy in electronic medical record-reported wait times to initial emergency physician evaluation and nursing staff requirement!

why hourly forecast is important?

12 hour forecasts are required to inform the Operations Department and in particular the Operational Delivery Unit (ODU) of future demand, particularly for the current and the next shift. The combination of forecast demand, incidents being attended, resource availability and delays at hospital provide information on the state of the unscheduled care system across Wales. Having this full picture enables the ODU to focus on the areas that require intervention to enable the most effective delivery of the service to the patients of Wales.

Forecasting hourly admissions is crucial for operational planning which  involves the short-term decision making related to the execution of the delivery process for various health care services such as Ambulatory ,Emergency, Surgical, Inpatient, Home and Residential.

- Adjusting the Schedule during  a day or a week because of unplanned events, such as emergency or walk-in patients, extended consultation times, and equipment
- Dynamic patient (re)assignment
- Staff rescheduling At the start of a shift, the staff sche- dule is reconsidered. Before and during the shift, the staff capacities may be adjusted to unpredicted demand fluctuations and staff absenteeism by using part-time, on- call nurses, staff overtime and voluntary absenteeism

what literature says about hourly forecasting in healthcare?

There exists a large literature on forecasting patient arrival, however the variation in arrivals remain unaccounted for. The forecasted hourly admissions at a typical A&E or Ambulance service show a mean absolute percentage error (MAPE) of 50%. Hourly forecasts are challenging because the noise caused by random variation may overshadow any pattern in the data. In this respect, forecasts of daily or monthly arrivals are likely easier but target decisions about staff allocation and the like, not the ongoing scheduling and rescheduling of how the available resources are divided among the patients in need of emergency services

limitations in current literature?

how we add to the literature?

@@

1. We develope a novel methodology to forecast hourly time series for healthcare services combining a distributio fit to each hour of day seperately and then adjust it using variables accounting for i) trend, ii) autocorrelation iii) temporal factors such as day of week, ... iv) special events such as public holidays, festive days and rugby and v) weather data.
2. We provide prbabilistic forecasts ...
3. We benchmark the forecasting performance of the proposed methodology against regression, prophet and TBATS using ....

The rest of the paper is organised as following: section \ref{lit} provides a brief overview of the use of temporal aggregation in time series forecasting. Section \ref{model} starts with ...

# Research background: Hourly forecasting in care services{#lit}

@@

\citet{hertzum2017forecasting} & Hourly ED patient arrivals and ED occupancy forecasting using calendar variables  1-3 & Regression, ARIMA, naive & MAPE & blabla\\ 
\citet{mccarthy2008challenge} & To develop methodology for predicting demand for ED services by characterizing ED arrivals  1-3 & Poisson regression & RMSE & blabla\\ 
\citet{morzuch2006forecasting} & hourly ED arrivals for 24 hours & Holt-Winters & MAPE & blabla\\ 
\citet{chase2012predicting} & ED volume & Logistic regression & RMSE & blabla\\ 
\citet{jones2009multivariate} & Demands for key resources in the ED and the inpatient hospital  & S, ARIMA & RMSE & blabla\\

@@

Linear regression, ARIMA, and naive models were used by @hertzum2017forecasting to investigate whether accurate hourly accident and emergency department patient arrivals and occupancy forecasts can be generated using calendar variables. Naive model was there for the purpose of comparison. @hertzum2017forecasting study shows that patient arrivals variation is larger across the hours of the day than across the days of the week and the months of the year. In term of hour of the day, patient arrivals peaked around noon. For days of the week, Monday is the busiest day while weekends are the quietest days. July-August are the month with the highest number of patient arrivals and January and February are the months with the lowest number of arrivals. The regression and ARIMA models perform similarly for all forecast interval in modeling patient arrivals. In modeling accident and emergency department occupancy, ARIMA outperform regression models. However, after all, the models of occupancy were less accurate than those arrivals. @hertzum2017forecasting mentioned that ARIMA models are among the most accurate models for accident and emergency department visits forecasting. Another interesting point is that the accuracy of accident and emergency department forecasting models decrease with the increasing forecast interval. Lastly, the accuracy of the forecasting model may possibly be increased with additional information added to the model.

@@

@@

```{r hourly-plot, fig.cap = 'Seasonal plot of ED attendance', fig.align='center',fig.height=9, fig.width= 6}
h2_hourly <- h2_hourly %>% filter(arrival_1h >="2015-01-01 00:00:00"&arrival_1h <"2019-01-01 00:00:00")
  subseriesplot <- h2_hourly %>% as_tibble() %>% mutate(day=factor(yday(arrival_1h)),year=factor(year(arrival_1h)),hour=hour(arrival_1h)) %>% 
    select(-arrival_1h)
time <-  format( seq.POSIXt(as.POSIXct(Sys.Date()), as.POSIXct(Sys.Date()+1), by = "1 hour"),
                 "%H:%M", tz="GMT")
time <- time[-length(time)]
p1 <- ggplot(data=subseriesplot,mapping = aes(x=hour ,y=n_attendance, colour=day))+
    geom_line(aes(group=interaction(day,year)))+
    geom_point()+
    theme_bw()+
    theme(legend.position = "")+
    labs(y="Number of visits in ED", x="")+
      scale_x_continuous(breaks=0:23, labels = time)+
  theme(axis.text.x = element_text(angle = 90))+
  theme(plot.title = element_text(size = 10))

p2 <- subseriesplot %>% 
  ggplot(mapping = aes(x=factor(hour) ,y=n_attendance))+
  geom_boxplot(aes(fill=year))+
  scale_color_brewer(palette = "Set1")+
  theme_bw()+
  theme(legend.position = "bottom", legend.box = "horizontal")+
  labs(fill="Year",y="Number of visits in ED", x="Hour of day")+
  scale_x_discrete(breaks=0:23, labels = time)+
  theme(axis.text.x = element_text(angle = 90))+
  theme(plot.title = element_text(size = 10))

h <- tibble(h=rep(1:24,1461 ))
ae_h_hour <- bind_cols(h2_hourly,h)
ae_h_hour <- ae_h_hour %>% as_tibble() %>% mutate(date=date(arrival_1h)) %>% select(h,date,n_attendance)
p <- ggplot(ae_h_hour, mapping = aes(x=factor(h), y=n_attendance))+
  geom_boxplot(outlier.colour = "red")+
  stat_summary(fun = quantile, geom="line",fun.args = list(probs = 0.05) , group= 1, aes(color= "#0072B2"), size = .3)+
  stat_summary(fun = quantile, geom="line",fun.args = list(probs = 0.25) , group= 1, aes(color= "#009E73"), size = .3)+
  stat_summary(fun = median, geom="line", group= 1, aes(color= "#CC79A7"), size = .5)+
  stat_summary(fun = quantile, geom="line",fun.args = list(probs = 0.75) , group= 1, aes(color= "#E69F00"), size = .3)+
  stat_summary(fun = quantile, geom="line",fun.args = list(probs = 0.95) , group= 1, aes(color= "#56B4E9"), size = .3)+
  stat_summary(fun = mean, geom="line", group= 1, aes(color= "#F0E442"), size = .5)

time <-  format( seq.POSIXt(as.POSIXct(Sys.Date()), as.POSIXct(Sys.Date()+1), by = "1 hour"),
                 "%H:%M", tz="GMT")
time <- time[-length(time)]

p_boxplot <- p+ theme_light()+
  labs(x="Hour of day", y="Number of visits in ED", color="")+ # Set line color to blue
  scale_color_identity(
                       breaks = c("#0072B2","#009E73","#CC79A7","#F0E442","#E69F00","#56B4E9"),
                       labels = c("5% quantile", "25% quantile","Median", "Mean","75% quantile","95% quantile"),
                       guide = "legend")+
  theme(legend.position = "bottom")+
  guides(color=guide_legend(nrow=1,byrow=TRUE))+
  scale_x_discrete(breaks=1:24,labels=time)+
  theme(axis.text.x = element_text(angle = 90))

p1/p2/p_boxplot
```

```{r 24hour, fig.cap= "Distrubution of ED attendance for each hour of day", fig.align='center',fig.width=6, fig.height=9, dependson="hourly-plot"}

hour.labs <- time
names(hour.labs) <- c(0:23)

h2_hourly %>% as_tibble() %>% mutate(Date=date(arrival_1h),Hour=hour(arrival_1h)) %>% 
  select(-arrival_1h) %>% ggplot(mapping=aes(x=n_attendance))+ 
geom_histogram(mapping=aes(y=stat(density)), binwidth = 1,fill="white",color = 'black')+
  geom_density(color="blue")+
scale_y_continuous(labels = scales::percent_format())+
facet_wrap(~Hour, scales = "free", ncol = 3,labeller = labeller(Hour = hour.labs))+
theme_bw()+
labs(x="Number of visits in ED", y="Percentage")+
theme(axis.text = element_text(size = 7))
```

### 2020-06-02 paper/paper.tex
add introduction and new graphs to data subsection

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[top=25mm, left=30mm, right=30mm, bottom=25mm,headsep=10mm, footskip=12mm]{geometry}
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Short-term hourly forecasting in urgent and emergency care},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Pandoc header
\usepackage{adjustbox, float}

@@

  \title{Short-term hourly forecasting in urgent and emergency care}
    \author[University1]{Author1\corref{1}}
   \ead{email1@example.com} 
    \author[University2]{Author2\corref{2}}
   \ead{email2@example.com} 
    \author[University3]{Author3\corref{2}}
   \ead{email3@example.com} 
      \address[University1]{Cardiff business school, 3 Colum Drive, CF10 3EU, Cardiff}
    \address[University2]{adress2}
    \address[University3]{adress3}
      \cortext[1]{Corresponding Author}
    \cortext[2]{Equal contribution}

@@

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  We measure the characteristics of the time series at the original level and at various levels of aggregation. We reveal how TA influences the time series characteristics.
\item
  We develope a model to discover the relationship between the original time series characteristics and the forecasting performance evaluation results to construct rules for method selection.
\item
  We determine for which type of time series characteristics, each TA forecasting strategy is preferable.
\end{enumerate}

The rest of the paper is organised as follwoing: section \ref{lit} provides a brief overview of the use of temporal aggregation in time series forecasting. Section \ref{model} starts with \ldots{}

\hypertarget{lit}{%
\section{Research background: Hourly short-term forecasting in emergency care}\label{lit}}

@@

Linear regression, ARIMA, and naive models were used by \citet{hertzum2017forecasting} to investigate whether accurate hourly accident and emergency department patient arrivals and occupancy forecasts can be generated using calendar variables. Naive model was there for the purpose of comparison. \citet{hertzum2017forecasting} study shows that patient arrivals variation is larger across the hours of the day than across the days of the week and the months of the year. In term of hour of the day, patient arrivals peaked around noon. For days of the week, Monday is the busiest day while weekends are the quietest days. July-August are the month with the highest number of patient arrivals and January and February are the months with the lowest number of arrivals. The regression and ARIMA models perform similarly for all forecast interval in modeling patient arrivals. In modeling accident and emergency department occupancy, ARIMA outperform regression models. However, afterall, the models of occupancy were less accurate than those arrivals. \citet{hertzum2017forecasting} mentioned that ARIMA models are among the most accurate models for accident and emergency department visits forecasting. Another interesting point is that the accuracy of accident and emergency department forecasting models decrease with the increasing forecast interval. Lastly, the accuracy of the forecasting model may possibly be increased with additional information added to the model.

@@

{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/hourly-plot-1}

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[top=25mm, left=30mm, right=30mm, bottom=25mm,headsep=10mm, footskip=12mm]{geometry}
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Short-term hourly forecasting for care services},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Pandoc header
\usepackage{adjustbox, float,lscape}

@@

  \title{Short-term hourly forecasting for care services}
    \author[University1]{Author1\corref{1}}
   \ead{email1@example.com} 
    \author[University2]{Author2\corref{2}}
   \ead{email2@example.com} 
    \author[University3]{Author3\corref{2}}
   \ead{email3@example.com} 
      \address[University1]{Cardiff business school, 3 Colum Drive, CF10 3EU, Cardiff}
    \address[University2]{adress2}
    \address[University3]{adress3}
      \cortext[1]{Corresponding Author}
    \cortext[2]{Equal contribution}

@@

why forecasting for urgent and emergency care is important?

Urgent and emergency healthcare is increasingly regarding patients as consumers of its service and aims to provide `consumer satisfaction'; particularly in the Accident and Emergency (A\&E) department and Ambulance services which have been viewed as the `shop window' of the hospital service. An accurate forecasting of the demand is essential in A\&E and Ambulance service trust to depict various courses of action that can result in massive savings in terms of patient lives.

Accurate forecasts contribute to better allocation of resources. Patient overcrowding in urgent and emergency care is a serious problem that causes challenging situations on patient flow . Also, it is related with increasing length of stay (LOS), low patient satisfaction about treatments on patients, number of patients left the ED without seen by medical staff , unexpected return visits to EDs , increasing health care costs and some other miscellaneous problems such as inaccuracy in electronic medical record-reported wait times to initial emergency physician evaluation and nursing staff requirement!

why hourly forecast is important?

12 hour forecasts are required to inform the Operations Department and in particular the Operational Delivery Unit (ODU) of future demand, particularly for the current and the next shift. The combination of forecast demand, incidents being attended, resource availability and delays at hospital provide information on the state of the unscheduled care system across Wales. Having this full picture enables the ODU to focus on the areas that require intervention to enable the most effective delivery of the service to the patients of Wales.

Forecasting hourly admissions is crucial for operational planning which involves the short-term decision making related to the execution of the delivery process for various health care services such as Ambulatory ,Emergency, Surgical, Inpatient, Home and Residential.

\begin{itemize}
\tightlist
\item
  Adjusting the Schedule during a day or a week because of unplanned events, such as emergency or walk-in patients, extended consultation times, and equipment
\item
  Dynamic patient (re)assignment
\item
  Staff rescheduling At the start of a shift, the staff sche- dule is reconsidered. Before and during the shift, the staff capacities may be adjusted to unpredicted demand fluctuations and staff absenteeism by using part-time, on- call nurses, staff overtime and voluntary absenteeism
\end{itemize}

what literature says about hourly forecasting in healthcare?

There exists a large literature on forecasting patient arrival, however the variation in arrivals remain unaccounted for. The forecasted hourly admissions at a typical A\&E or Ambulance service show a mean absolute percentage error (MAPE) of 50\%. Hourly forecasts are challenging because the noise caused by random variation may overshadow any pattern in the data. In this respect, forecasts of daily or monthly arrivals are likely easier but target decisions about staff allocation and the like, not the ongoing scheduling and rescheduling of how the available resources are divided among the patients in need of emergency services

limitations in current literature?

how we add to the literature?

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  We develope a novel methodology to forecast hourly time series for healthcare services combining a distributio fit to each hour of day seperately and then adjust it using variables accounting for i) trend, ii) autocorrelation iii) temporal factors such as day of week, \ldots{} iv) special events such as public holidays, festive days and rugby and v) weather data.
\item
  We provide prbabilistic forecasts \ldots{}
\item
  We benchmark the forecasting performance of the proposed methodology against regression, prophet and TBATS using \ldots{}.
\end{enumerate}

The rest of the paper is organised as following: section \ref{lit} provides a brief overview of the use of temporal aggregation in time series forecasting. Section \ref{model} starts with \ldots{}

\hypertarget{lit}{%
\section{Research background: Hourly forecasting in care services}\label{lit}}

@@

Linear regression, ARIMA, and naive models were used by \citet{hertzum2017forecasting} to investigate whether accurate hourly accident and emergency department patient arrivals and occupancy forecasts can be generated using calendar variables. Naive model was there for the purpose of comparison. \citet{hertzum2017forecasting} study shows that patient arrivals variation is larger across the hours of the day than across the days of the week and the months of the year. In term of hour of the day, patient arrivals peaked around noon. For days of the week, Monday is the busiest day while weekends are the quietest days. July-August are the month with the highest number of patient arrivals and January and February are the months with the lowest number of arrivals. The regression and ARIMA models perform similarly for all forecast interval in modeling patient arrivals. In modeling accident and emergency department occupancy, ARIMA outperform regression models. However, after all, the models of occupancy were less accurate than those arrivals. \citet{hertzum2017forecasting} mentioned that ARIMA models are among the most accurate models for accident and emergency department visits forecasting. Another interesting point is that the accuracy of accident and emergency department forecasting models decrease with the increasing forecast interval. Lastly, the accuracy of the forecasting model may possibly be increased with additional information added to the model.

@@

{\centering \includegraphics{paper_files/figure-latex/hourly-plot-1}
